Multi-resolution Neural Field Estimation

### 2011-03-05 Multiresolution_framework/ltx/manuscript.tex
some edits to Mike comments

where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ is denoted $v_t\left(\mathbf r\right)$. Synaptic dynamics are included through the parameter $\xi=\left(1-\frac{ T_s}{\tau}\right)$, where $\tau$ is the synaptic time constant and $T_s$ is the sampling time. The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the \mike{connectivity} kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977}. The term $e_t(\mathbf r)$ is an i.i.d. disturbance, such that $e_t(\mathbf{s})\sim\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$\mike{:}\cut{, where $\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$  denotes} a zero mean spatial Gaussian process with covariance function $\eta(\mathbf{r}-\mathbf{r'})$~\cite{Rasmussen2005}. The firing rate of the presynaptic neurons is related to the post-synaptic membrane potential by the activation function $f(v_t(\mathbf{r})) = \varsigma v_t(\mathbf{r})$. A linear form of the activation function was chosen to construct the estimator due to high computational demands when using a nonlinear (sigmoidal) activation function. The observation equation describing the electrophysiological recordings is given by \mike{should the $r_n$ be $r_{n_y}$ below?} 

@@

	y_t(\mathbf{r}_n) = \int_{\Omega} { m\left(\mathbf{r}_{n}-\mathbf{r}'\right) v_t\left(\mathbf{r}'\right) \, d\mathbf{r}'} + \boldsymbol\epsilon_t, 

@@

where $m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right)$ is the observation kernel and the observation noise is an i.i.d. \mike{Gaussian} white noise process and $\boldsymbol{\epsilon}_{t}\sim \mathcal{N}\left(\mathbf{0},\mathbf{\Sigma}_{\epsilon}\right)$. % $\mathbf{y}_{t} = [y_t(\mathbf{r}_1) y_t(\mathbf{r}_2) \cdots y_t(\mathbf{r}_{n_y})]^\top$, compiled at $n_{y}$ spatial locations at time $t$ via the observation kernel , is corrupted by The superscript $\top$ denotes the transpose operator.

@@

Eqs.\eqref{eq:KernelExpansion} and \eqref{eq:FieldExpansion} are infinite series expansions and must be truncated at some level $j$. The field and the connectivity kernel are approximated by
\begin{align}
	w\left(r-r'\right) &\approx \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right) 
	\label{eq:KernelFiniteExpansion} \\
	v_t\left(r\right) &\approx \boldsymbol\mu\left(r\right)^\top\mathbf{x}_t,
	\label{eq:FieldFiniteExpansion}
\end{align}
where the unknown parameter and state vectors, $\boldsymbol\theta \in \mathbb{R}^{n_{\theta}}$ and $\mathbf{x}_t \in \mathbb{R}^{n_x}$, are defined as 
\begin{align}
\boldsymbol\theta^\top &=\left[ \boldsymbol\alpha_{j_0}^\top \quad \boldsymbol\beta_{j_0}^\top \quad \boldsymbol\beta_{j_0+1}^\top \cdots \boldsymbol\beta_{j}^\top\right] 
\label{KernelWeights} \\
\mathbf{x}_{t}^\top &=\left[\mathbf{x}_{t,j_{0}}^\top \quad  \check{\mathbf{x}}_{t,j_{0}}^\top \quad  \check{\mathbf{x}}_{t,j_{0}+1}^\top \cdots \check{\mathbf{x}}_{t,j}^\top\right].
\label{FieldWeights}
\end{align}
\mike{I think it would be much clearer to simply define these vectors as vectors, rather than hoping that the reader will make the same mapping from the sets defined here to the vectors you actually use.} The kernel and field approximation coefficient vectors, $\boldsymbol \alpha_{j_0}$ and $\mathbf{x}_{t,j_{0}}$, contain all the coefficients $\left\lbrace\alpha_{j_0, l}:l \in \mathbb{Z} \right\rbrace $ and $\left\lbrace x_{t,j_0, l}: l \in \mathbb{Z}\right\rbrace$, respectively. Similarly, the kernel and the field detail coefficient vectors, $\boldsymbol\beta_{j}$ and $\check{\mathbf{x}}_{t,j}$ contain all the coefficients $\left\lbrace \beta_{j,l} :l \in \mathbb{Z}\right\rbrace$ and $\left\lbrace  \check x_{t,j, l}:l \in \mathbb{Z}\right\rbrace$ respectively.

The vectors of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$ respectively, are defined by
\begin{align}
 \boldsymbol\lambda^\top & (r-r')=\left[ \boldsymbol\phi_{j_0}^\top(r-r') \quad \boldsymbol\psi_{j_0}^\top(r-r') \quad \boldsymbol\psi_{j_0+1}^\top(r-r') \right. \nonumber \\
&\left. \cdots \quad \boldsymbol\psi_{j}^\top(r-r')\right] \label{KernelBasisVector} \\
 \boldsymbol\mu^\top & (r)=\left[ \boldsymbol\phi_{j_0}^\top(r) \quad \boldsymbol\psi_{j_0}^\top(r) \quad \boldsymbol\psi_{j_0+1}^\top(r) \cdots \boldsymbol\psi_{j}^\top(r) \right]. 
\label{FieldBasisVector}
\end{align}
\mike{is it worth defining $s=r-r'$ somewhere and using that throughout? It may save some lines!} The vectors in Eqs. (\ref{KernelBasisVector}) and (\ref{FieldBasisVector}) are constructed in the same manner as the vectors in Eqs. (\ref{KernelWeights}) and (\ref{FieldWeights}).

To derive the state-space model, we first define the matrices \mike{bold $r$s here and everywhere (or nowhere)?}
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\mathbf{x}_t 
	\, dr'dr}
\end{align}
Eqs.~\eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} are then substituted into Eq. \eqref{eq:DiscreteTimeModel}, then we \mike{do you mean pre-multiply? - I'm not sure what cross multiply means here} cross-multiply by $\boldsymbol\mu\left(\mathbf r\right)$ and integrate over the space giving
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1}= 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= (T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I),
\end{align}
where $\mathbf I$ is the identity matrix. The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed white noise process with covariance (see \cite{Scerri2009} for proof)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into equation \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}

@@

A direct consequence of Eq.~\eqref{eq:BsplineConvolution} is \dean{the inner product between two B-splines is another translated B-spline, such that}

>>>>

where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ is denoted $v_t\left(\mathbf r\right)$. Synaptic dynamics are included through the parameter $\xi=\left(1-\frac{ T_s}{\tau}\right)$, where $\tau$ is the synaptic time constant and $T_s$ is the sampling time. The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the connectivity kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977}. The term $e_t(\mathbf r)$ is an i.i.d. disturbance, such that $e_t(\mathbf{s})\sim\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$: a zero mean spatial Gaussian process with covariance function $\eta(\mathbf{r}-\mathbf{r'})$~\cite{Rasmussen2005}. The firing rate of the presynaptic neurons is related to the post-synaptic membrane potential by the activation function $f(v_t(\mathbf{r})) = \varsigma v_t(\mathbf{r})$. A linear form of the activation function was chosen to construct the estimator due to high computational demands when using a nonlinear (sigmoidal) activation function. The observation equation describing the electrophysiological recordings is given by

@@

	y_t(\mathbf{r}_{n_y}) = \int_{\Omega} { m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right) v_t\left(\mathbf{r}'\right) \, d\mathbf{r}'} + \boldsymbol\epsilon_t(\mathbf{r}_{n_y}), 

@@

where $m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right)$ is the observation kernel at location $\mathbf{r}_{n_y}$ and the observation noise is an i.i.d. Gaussian white noise process and $\boldsymbol{\epsilon}_{t}\sim \mathcal{N}\left(\mathbf{0},\mathbf{\Sigma}_{\epsilon}\right)$. % $\mathbf{y}_{t} = [y_t(\mathbf{r}_1) y_t(\mathbf{r}_2) \cdots y_t(\mathbf{r}_{n_y})]^\top$, compiled at $n_{y}$ spatial locations at time $t$ via the observation kernel , is corrupted by The superscript $\top$ denotes the transpose operator.

@@

Eqs.\eqref{eq:KernelExpansion} and \eqref{eq:FieldExpansion} are infinite series expansions and must be truncated at some level $j$. The field and the connectivity kernel are approximated by
\begin{align}
	w\left(r-r'\right) &\approx \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right) 
	\label{eq:KernelFiniteExpansion} \\
	v_t\left(r\right) &\approx \boldsymbol\mu\left(r\right)^\top\mathbf{x}_t,
	\label{eq:FieldFiniteExpansion}
\end{align}
where the unknown parameter and state vectors, $\boldsymbol\theta \in \mathbb{R}^{n_{\theta}}$ and $\mathbf{x}_t \in \mathbb{R}^{n_x}$, are defined as 
\begin{align}
\boldsymbol\theta^\top &=\left[ \boldsymbol\alpha_{j_0}^\top \quad \boldsymbol\beta_{j_0}^\top \quad \boldsymbol\beta_{j_0+1}^\top \cdots \boldsymbol\beta_{j}^\top\right] 
\label{KernelWeights} \\
\mathbf{x}_{t}^\top &=\left[\mathbf{x}_{t,j_{0}}^\top \quad  \check{\mathbf{x}}_{t,j_{0}}^\top \quad  \check{\mathbf{x}}_{t,j_{0}+1}^\top \cdots \check{\mathbf{x}}_{t,j}^\top\right].
\label{FieldWeights}
\end{align}
The kernel and field approximation coefficient vectors, $\boldsymbol \alpha_{j_0}$ and $\mathbf{x}_{t,j_{0}}$, contain all the coefficients $\left\lbrace\alpha_{j_0, l}:l \in \mathbb{Z} \right\rbrace $ and $\left\lbrace x_{t,j_0, l}: l \in \mathbb{Z}\right\rbrace$, respectively. Similarly, the kernel and the field detail coefficient vectors, $\boldsymbol\beta_{j}$ and $\check{\mathbf{x}}_{t,j}$ contain all the coefficients $\left\lbrace \beta_{j,l} :l \in \mathbb{Z}\right\rbrace$ and $\left\lbrace  \check x_{t,j, l}:l \in \mathbb{Z}\right\rbrace$ respectively.

The vectors of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$ respectively, are defined by
\begin{align}
 \boldsymbol\lambda^\top & (r-r')=\left[ \boldsymbol\phi_{j_0}^\top(r-r') \quad \boldsymbol\psi_{j_0}^\top(r-r') \quad \boldsymbol\psi_{j_0+1}^\top(r-r') \right. \nonumber \\
&\left. \cdots \quad \boldsymbol\psi_{j}^\top(r-r')\right] \label{KernelBasisVector} \\
 \boldsymbol\mu^\top & (r)=\left[ \boldsymbol\phi_{j_0}^\top(r) \quad \boldsymbol\psi_{j_0}^\top(r) \quad \boldsymbol\psi_{j_0+1}^\top(r) \cdots \boldsymbol\psi_{j}^\top(r) \right]. 
\label{FieldBasisVector}
\end{align}
The vectors in Eqs. (\ref{KernelBasisVector}) and (\ref{FieldBasisVector}) are constructed in the same manner as the vectors in Eqs. (\ref{KernelWeights}) and (\ref{FieldWeights}).

To derive the state-space model, we first define the matrices \mike{bold $r$s here and everywhere (or nowhere)?}
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\mathbf{x}_t 
	\, dr'dr}
\end{align}
Eqs.~\eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} are then substituted into Eq. \eqref{eq:DiscreteTimeModel}, then we pre-multiply by $\boldsymbol\mu\left(\mathbf r\right)$ and integrate over the space giving
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1}= 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= (T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I),
\end{align}
where $\mathbf I$ is the identity matrix. The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed white noise process with covariance (see \cite{Scerri2009} for proof)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into equation \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}

@@

A direct consequence of Eq.~\eqref{eq:BsplineConvolution} is the inner product between two B-splines is another translated B-spline, such that
